---
title: "P8105 Final Project Data Cleaning"
author: "Yuki Low yl5503, Lauren lel2176, Trina eih2108, Arthur as7123"
date: "11/20/2023"
output:
  html_document:
    df_print: paged
---

```{r, message = FALSE}
library(tidyverse)
library(lubridate)
library(readr) 
library("ggplot2") 
library("dplyr")
library(xts)
library("lubridate")
library("RColorBrewer")
library("ggthemes")
library("gridExtra")
library("leaflet")
library("highcharter")
library(scales)
library(leaflet.extras)
```

```{r, warning = FALSE}
rats_raw <- read.csv("./Rat_Sightings.csv", na = c("", "NA", "N/A", "Unspecified")) %>%
  janitor::clean_names() %>% 
  mutate(created_date = mdy_hms(created_date)) %>%
  mutate(sighting_year = year(created_date),
         sighting_month_num = month(created_date),
         sighting_month = month(created_date, label = TRUE, abbr = FALSE),
         sighting_day = day(created_date),
         sighting_weekday = wday(created_date, label = TRUE, abbr = FALSE)) 

rats_raw

```

Overall trend in rats 
```{r}
overall <- rats_raw %>% 
  group_by(sighting_year, sighting_month_num, sighting_day) %>% 
  summarize(count = n()) %>% 
  mutate(date = as.Date(paste(sighting_year, sighting_month_num, sighting_day, sep = "-")))

time_series = xts(overall$count , order.by= overall$date)

hchart(time_series, name = "Rat Sightings") %>% 
  hc_add_theme(hc_theme_darkunica()) %>%
  hc_credits(enabled = TRUE, text = "Sources: City of New York", style = list(fontSize = "12px")) %>%
  hc_title(text = "Time Series of NYC Rat Sightings") %>%
  hc_legend(enabled = TRUE)
```

There seems to be an increasing trend of rat sightings in NYC from 2009 to 2023. 

Trend in rats by borough over time 
```{r}
borough_over_time <- rats_raw %>% 
  group_by(sighting_year, sighting_month_num, sighting_day, borough) %>% 
  summarize(count = n()) %>% 
  mutate(date = as.Date(paste(sighting_year, sighting_month_num, sighting_day, sep = "-"))) %>% 
  filter(!is.na(borough))

ggplot(borough_over_time, aes(x = date, y = count, color = borough)) +
  geom_line() +
  labs(title = "Entries Over Time by Borough",
       x = "Year",
       y = "Count") +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  theme_minimal()
```
From this graph, we can see that Brooklyn seems to have the most rat sightings in NYC. However, we are unable to see the trend in rat sightings in the Bronx. 


```{r}
borough_over_time_wo_brooklyn <- rats_raw %>% 
  filter(borough != "BROOKLYN") %>% 
  filter(borough != "MANHATTAN") %>% 
  group_by(sighting_year, sighting_month_num, sighting_day, borough) %>% 
  summarize(count = n()) %>% 
  mutate(date = as.Date(paste(sighting_year, sighting_month_num, sighting_day, sep = "-"))) %>% 
  filter(!is.na(borough))

borough_over_time_wo_brooklyn

ggplot(borough_over_time_wo_brooklyn, aes(x = date, y = count, color = borough)) +
  geom_line() +
  labs(title = "Entries Over Time by Borough",
       x = "Year",
       y = "Count") +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  theme_minimal()
```
It seems like the rat sightings in Bronx and Queens are comparable but further EDA is needed to see where the Bronx stands in terms of rat sightings. 

Count of rats by year 
```{r}
by_year <- rats_raw %>% 
  group_by(sighting_year) %>% 
  count() %>% 
  ggplot(aes(x = sighting_year, y = n, fill = n)) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme(legend.position ='none',axis.title = element_text(),axis.text.x = element_text(size = 12)) +
  xlab("Year") + 
  ylab("Count") +
  geom_text(aes(label = n), vjust = -0.1, size = 3.75) +
  ggtitle('Count of Rat Sightings through the Years') + 
  scale_fill_gradientn(name = '',colours = rev(brewer.pal(10,'Spectral'))) 

by_year
```
There was a huge jump in total number of rat sightings from 2020 to 2021. 

Counts of rats by month 
```{r}
by_month <- rats_raw %>% 
  group_by(sighting_month) %>% 
  count() %>% 
  ggplot(aes(x = sighting_month, y = n, fill = n)) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme(legend.position ='none',axis.title = element_text(),axis.text.x = element_text(size = 9)) +
  xlab("Month") + 
  ylab("Count") +
  geom_text(aes(label = n), vjust = -0.1, size = 3.75) +
  ggtitle('Count of Rat Sightings by Month') + 
  scale_fill_gradientn(name = '',colours = rev(brewer.pal(10,'Spectral'))) 

by_month
```
It seems like there are the most rat sightings in the warmer months with the peak being at July. 

Counts of rat by day of the week 
```{r}
by_day <- rats_raw %>% 
  group_by(sighting_weekday) %>% 
  count() %>% 
  ggplot(aes(x = sighting_weekday, y = n, fill = n)) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme(legend.position ='none',axis.title = element_text(),axis.text.x = element_text(size = 12)) +
  xlab("Weekday") + 
  ylab("Count") +
  geom_text(aes(label = n), vjust = -0.1, size = 4) +
  ggtitle('Count of Rat Sightings by Day of Week') + 
  scale_fill_gradientn(name = '',colours = rev(brewer.pal(10,'Spectral'))) 

by_day
```
There are more rat sighings reported during the weekdays compared to weekends. 

```{r}
by_location_type <- rats_raw %>% 
  group_by(location_type) %>% 
  filter(location_type != "Other (Explain Below)") %>% 
  count() %>% 
  arrange(desc(n)) %>%
  head(30)

by_location_type

ggplot(by_location_type, aes(x = location_type, y = n)) +
  geom_histogram(stat = "identity", position = "dodge") +
  labs(title = "Sightings by Location Type",
       x = "Location Type",
       y = "Count") +
  theme_minimal() + 
  coord_flip() + 
  geom_text(aes(label = n), hjust = -0.01, size = 3) 

```
```{r split = FALSE, fig.align = 'default', warning = FALSE, out.width="100%"}
## Overall Sightings Map and Heat Map

top = 40.917577 # north lat
left = -74.259090 # west long
right = -73.700272 # east long
bottom =  40.477399 # south lat


nyc = rats_raw %>%
  filter(latitude >= bottom) %>%
  filter ( latitude <= top) %>%
  filter( longitude >= left ) %>%
  filter(longitude <= right)

center_lon = median(nyc$longitude,na.rm = TRUE)
center_lat = median(nyc$latitude,na.rm = TRUE)

count = nyc %>%
  group_by(location) %>%
  count()

count 

nyc = merge(nyc, count, by = "location")

factpal = colorFactor("blue", nyc$n)

nyc %>%
leaflet() %>% 
  addProviderTiles("Esri.NatGeoWorldMap") %>%
  addCircles(lng = ~longitude, lat = ~latitude)  %>%
  setView(lng=center_lon, lat=center_lat,zoom = 10) 

nyc %>%
  leaflet() %>%
  addProviderTiles("Esri.NatGeoWorldMap") %>%
  addHeatmap(lng = ~longitude, lat = ~latitude, intensity = ~(nyc$n), blur = 20, max = 0.05, radius = 15) %>%
  setView(lng=center_lon, lat=center_lat,zoom = 10)

```
